Synthesizing Symmetric Lenses

### 2018-07-08 papers/SymmetricLenses/main.tex
add this work

In summary, our main contributions are:
\begin{itemize}
\item We define the space of simple symmetric lenses, which lies between
  asymmetric lenses and full symmetric lenses and has the useful property that
  its members don't have any internal state (\S\ref{sec:overview}). 
\item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case),
using a novel application of stochastic regular expressions to guide the search process.  In particular, we
show that the RE star semi-ring equivalences remain valid when their semantics
are enriched by distributions (\S\ref{sec:synthesis}).
\item We extend the Boomerang implementation with simple symmetric lenses,
  showing that they can easily be combined with a rich variety of other lens
  extensions including quotients and matching lenses. We evaluate our
  implementation over a set of benchmarks drawn from data cleaning, information
  extraction, and file synchronization tasks and demonstrate that we can
  synthesize all of them in under half a minute (\S\ref{sec:evaluation}).
%
  We also analyze the gap between simple and full symmetric lenses
  empirically and argue that simple ones suffice for many applications.
\item We show that the class of simple lenses can be characterized semantically
  as the subset of full symmetric lenses for which a simple ``forgetfulness''
  property holds (\S\ref{sec:relationship}).
\end{itemize}

@@

In this section, we give an overview of simple symmetric lenses and our
corresponding synthesis algorithm using an example drawn from a hypothetical
U.S. company. In this company, management and human resources (HR) store
information about employees in separate files: management records the names of
employees and their salaries while HR stores the names of employees and their
health insurance providers. Figure~\ref{fig:minimized-representations} gives
examples of the two types of files and regular expressions describing the two
formats. The company uses simple symmetric lenses to keep these files
synchronized. If management adds a new employee, ``Chris Roe: 32500'', a lens
adds the corresponding entry ``Chris Roe, UNK'' to HR's file. The sentinel value
``UNK'' represents the fact that the employee's insurance company is currently
unknown. A similar update happens if HR adds a new employee before management,
in which case the sentinel value ``unk'' reflects the unknown salary information
Furthermore, if HR corrects an error in an employee's name, say changing ``John
Public'' to ``Jon Public'', the lens mirrors that name change in management's
file. Not all the data is mirrored, however. The management file is not updated
in response to insurance changes, and the HR file is oblivious to salary
changes. Simple symmetric lenses are appropriate for keeping these two files
synchronized because the files have a mix of shared and unshared information.

@@

\subsection{Simple Symmetric Lenses}
A simple symmetric lens between formats described by regular expression types
$\BRegex$ and $\BRegexAlt$\footnote{We use nonterminals \BRegex and \BRegexAlt
  for regular expressions, using \Regex and \RegexAlt for stochastic regular
  expressions.} is defined by four functions.\bcp{I think a picture would be
  helpful. I can probably make one pretty easily by tweaking the one from the
  original symmetric paper.} \afm{That would be awesome.}

@@

\begin{gather}
 \tag{1}
 \mbox{$\CreateR \OfType \BRegex \rightarrow \BRegexAlt$}\\
 \tag{2}
 \mbox{$\CreateL \OfType \BRegexAlt \rightarrow \BRegex$}\\
 \tag{3}
 \mbox{$\PutR \OfType \BRegex \rightarrow \BRegexAlt \rightarrow \BRegexAlt$}\\
 \tag{4}
 \mbox{$\PutL \OfType \BRegexAlt \rightarrow \BRegex \rightarrow \BRegex$}
\end{gather}

\noindent
that collectively satisfy four ``round-tripping'' laws: 
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{s})}{s} = s\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{t})}{t} = t\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{s}{t})}{s} = s\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{t}{s})}{t} = t
\end{gather}
\noindent
The two create functions provide default values for times when
data appears on one side but not the other (\EG, the ``unk'' salary
entry added to the management file when HR inserts a new employee).
The two put functions propagate edits from one format to the
other by combining the change from one file with the previous value
from the other file. The roundtripping laws guarantee that identify edits to one file do
not change the data in the other file.

@@

We use regular expressions to denote the types of data formats, and we
impose a typing discipline on our lenses.  Specifically, if 
$\BRegex$ and $\BRegexAlt$ are regular expressions, then the notation
\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that $\Lens$
is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
$\LanguageOf{\BRegexAlt}$. 
When clear from context, we use the term \textit{lens} to denote a
simple symmetric lens.

\subsection{Simple Symmetric Lens Combinators} 
Generally, we do not expect programmers to define simple symmetric lenses by
writing the four functions by hand and showing they satisfy the
roundtripping laws.  Instead, we provide simple symmetric lens
combinators to build up the requisite lenses from simpler ones.
We briefly explain a selection of these combinators by using them to
define a lens between 
the two employee data formats in Figure~\ref{fig:minimized-representations}. 
The finished lens, which appears in Figure~\ref{fig:example_lens}, has
the type 
\lstinline{emp_salaries} $\Leftrightarrow$ \lstinline{emp_insurance}, where
\lstinline{emp_salaries} and \lstinline{emp_insurance} are regular
expressions defined in Figure~\ref{fig:minimized-representations}.

@@

In contrast to the identity lens, the \Disconnect lens, leaves data
unconnected. The disconnect lens takes four arguments: two regular expressions
and two strings. The regular expressions specify the formats of the
corresponding data on the two sides while the strings provide default values.
On creates, the input values are thrown away, and default values are returned ($\CreateROf{\String} =
\lstinline{"UNK"}$), and on puts, the second argument is used, and the first is
thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$).  In
our extended example, the \lstinline{salary} and
\lstinline{company} fields are independent, so we use the disconnect lens to relate them. The defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disconnect(salary, company, "unk", "UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
The insert \lstinline{ins} and delete \lstinline{del} lenses are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source and target formats,
respectively.
\begin{lstlisting}
 ins($t$) = disconnect("", $t$, "", $t$)
 del($s$) = disconnect($s$, "", $s$, "")
\end{lstlisting}

@@

We provide typing rules for each of the simple symmetric combinators, most of
which are syntax directed. However, regular expressions have a number of
equivalences on them, and sometimes we must convert one regular expression type
to another, equivalent regular expression type. Our typing rules include type
equivalence, so if a lens is typed between two regular expressions, it is also
typed between two equivalent ones. The flexibility this type equivalence rule
provides is useful when a different representation of a regular expression would
facilitate synthesis.

@@

While our example is relatively simple for demonstration, writing such lenses
for real-world formats is quite difficult, as these formats can be large and
transform data in complex and unintuitive ways. To alleviate the difficulty of
manually writing simple symmetric lenses, we wish to synthesize them instead.
Given a pair of regular expressions and a set of input-output examples, we want
to find a simple symmetric lens that is typed by the regular expression pair and
satisfies the input/output examples. We call such a lens a \emph{satisfying
  lens}. In our running example, we wish to synthesize a lens between
\lstinline{emp_salaries} and \lstinline{emp_insurance}, using as an input-output
example the data in Figure~\ref{fig:minimized-representations}. The algorithm
uses such examples to find suitable defaults for \Disconnect lenses and to
figure out how to match data that could be synchronised in multiple ways.

A challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disconnect(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is another valid lens (where \lstinline{rep($\BRegex$)} is a string in the
language of $\BRegex$). Furthermore, adding the input-output example in
Figure~\ref{fig:minimized-representations} 
helps, but only a little, because it requires many examples to rule
out all the possible occurrences of nested \Disconnect lenses,
particularly in complex formats. 
%\ksf{I don't understand the following.  We shouldn't explain the issue in
%terms of \PutR and \PutL as we are focused on the combinators rather
%than the raw functions.}
%The default strings given to \Disconnect merely need to be the strings
%in the input-output examples, and the lens will still satisfy the
%specification.  While the disconnect lens is prohibited by two
%examples, or by a carefully chosen input-output example on \PutR
%or \PutL, this problem continues presenting itself for the sublenses,
%and is highlighted even more on larger formats.
While we could ask programmers to provide detailed and precise specifications
for their synthesis tasks, we would prefer an algorithm that can intelligently
choose a lens despite the underspecification. Instead of merely finding
\emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
likely to please the user.

But what is a likely satisfying lens? We propose that a satisfying
lens is more likely if it connects more data between the two
formats, \IE, by using an identity rather than a disconnect lens.  The
likeliest satisfying lenses would connect as much data as
possible. More formally, we propose that the likelihood of a
satisfying lens is the expected number of bits required to recover
data in one format from synchronized data in the other.  Higher
likelihoods correspond to fewer numbers of bits. In more
detail, we say that two strings $s$ and $t$ are \emph{synchronized
according to lens $l$} if $l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$.\footnote{
The notation $l.\PutR$ extracts the \PutR function from simple
symmetric lens $l$; similarly for \PutL.}  We say that we
can \emph{recover} $s$ from $t$ using bits $b$ and lens $l$ if we can
reconstruct $s$ from $t$, $b$, and $l$.  For example, given
the \IdentityLens{} lens, we can recover $s$ from $t$ using no bits
because in this case, $s$ is just $t$.  In contrast, given
the \Disconnect{} lens, we need enough bits to fully encode $s$ in
order to recover it from $t$ because $t$ contains no useful
information.  This calculation shows that if both \IdentityLens{}
and \Disconnect{} are satisfying lenses, then \IdentityLens{} is more
likely.

The expected number of bits required to recover data is a
well-known concept in information theory: it is the \emph{entropy} of
the data. Calculating the 
entropy of data requires a probability distribution over the
space of possible values for the data.  Specifically,
given a set $S$ and a probability distribution over that set $P
\OfType S \rightarrow \Reals$, the entropy of the $(S,P)$ pair is $-\Sigma_{s\in
  S}P(s)\Log_2P(s)$.

In our setting, we already have a way of expressing sets of data:
regular expressions.  What we need is a way to express probability
distributions over that data.
To that end, we adopt \emph{stochastic regular expressions}\cite{?} (SREs), 
which are regular expressions where each operator is annotated with a
probability.  Stochastic regular expressions
allow us to jointly specify the allowable data and give a probability
distribution over it.  From this information, we can calculate the
entropy of the SRE.

This infrastructure gives us what we need to find likely satisfying lenses from
a collection of satisfying lenses. All such lenses must synchronize data between
the same two SREs $\Regex$ and $\RegexAlt$. For any such lens $\Lens$, we can
calculate the expected number of bits required to recover a string $s$ in
$\LanguageOf{\Regex}$ from a synchronized string $t$ in
$\LanguageOf{\RegexAlt}$. This expected number of bits is just the conditional
entropy of $\Regex$ given $\RegexAlt$ and $\Lens$. To ensure symmetry, the
likelihood we assign to $\Lens$ is the sum of the conditional entropy of
$\Regex$ given $\RegexAlt$ and $\Lens$ and the conditional entropy of
$\RegexAlt$ given $\Regex$ and $\Lens$.

@@

Now that we have a way to calculate likelihood, we need a way to search for
it.
Previous work on synthesis~\cite{?,?,?} has shown
that we can use types to dramatically restrict the search space and
improve the effectiveness of synthesis.  In our setting, types are
pairs of stochastic regular expressions, which come with a rich
algebraic structure that makes each regular expression semantically
equal to infinitely many others.  To tame that search space and
leverage the benefits of type-directed synthesis, we split our
algorithm into two communicating synthesizers following earlier
work~\cite{quotient-lens-synthesis}.  The first
synthesizer, \RXSearch, navigates the space of semantically equivalent
regular expressions by iteratively applying rewrite rules that
preserve both the semantics and the probability
distributions of those stochastic regular expressions. \RXSearch
ranks pairs of stochastic regular expressions 
by the number of rewrite rule applications required to obtain the pair
from the stochastic regular expressions given as input. It passes the
pairs off to the second synthesizer, \GreedySynth, in rank order with
the smallest first.

At a high level, \GreedySynth looks for lenses between a given pair of
stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the
highest likelihood by performing a type-directed search. 
In more detail, \GreedySynth first uses type equalities based on semiring 
equivalences (ignoring the star semiring equivalences that are managed
by \RXSearch) to normalize the
stochastic regular expressions provided by \RXSearch into 
\emph{stochastic n-ary DNF regular expressions}.  
Then, \GreedySynth uses the syntax of these n-ary DNF regular
expressions to find highly normalized lenses, in a
form we call \emph{simple symmetric n-ary DNF lenses}.  These lenses include
neither a composition operator nor a type equivalence rule.  These
restrictions mean that each simple symmetric n-ary DNF lens is typed
by a single pair of stochastic n-ary DNF regular expressions and 
\GreedySynth's search space for a given pair of regular expressions is
finite.  The variations that \GreedySynth 
considers include choosing between \IdentityLens{} and  \Disconnect
lenses and deciding how to match multiple occurrences of the same kind
of data, \EG, the two occurrences of \lstinline{name} in the
management and HR data formats (we want to map first names to first
names and last names to last names). 
Finally, \GreedySynth converts this lens 
back into a simple symmetric lens by converting the n-ary forms into
the binary forms provided in the surface language.

@@

\section{Stochastic Regular Expressions}
\label{sec:srx}
To characterize the likely lenses, we must compute the expected number of bits
needed to recover a string in one data source from a synchronized string in the
other data source. To do this, we must first develop a probabilistic langauge
model for our language. We do this with \emph{stochastic regular expressions},
regular expressions annotated with probability information. With this, we can
jointly express a language and a probability distribution over that language. In
the following grammar, $\String$ ranges over strings, and $\Probability$ ranges
over real numbers between 0 and 1, exclusive.

@@

In the stochastic regular expression
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
represents the likelihood the string is in $\Regex_1$. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability how
much more likely having a string involving more than $n$ iterations is than
having a string involving $n$ iterations. The semantics of SREs is defined by
the function $P$ below.

@@

\subsection{Stochastic Regular Expression Equivalences}
$\RXSearch$ needs to be able to find equivalent stochastic regular expressions.
However, as of yet, there is no syntactic means to reason about stochastic
regular expression equivalence. We extend the \emph{star-semiring} equivalences
to stochastic regular expressions.

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  The proofs of this theorem, and subsequent theorems in this paper, appears in
  the auxiliary materials accompanying this submission.
\end{proof}

\subsection{Stochastic Regular Expression Entropy}
The entropy of a data source $S$ is the expected number of bits required to
describe a random element in $S$ ($-\Sigma_{s\in S}P(s)\Log_2P(s)$). The entropy
of a stochastic regular expression can be computed directly from its syntax,
when the stochastic regular expression is unambiguous, and contains no nonempty
components.
\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{s}$
    & =
    & 1\\

@@

Without unambiguity constraints it is not easy to calculate the entropy. For
example, consider the SRE $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$.
Using the above computation, this SRE has an entropy of $1$, but the entropy
should be $0$. Similar issues occur when trying to find the entropy when
$\emptyset$ is a subterm, so we do not define the entropy on $\emptyset$.

In our work, we do not find this to be a problem, we require unambiguous regular
expressions as input to our synthesis procedure, and we can easily preprocess
emptysets out of the stochastic regular expressions that are themselves nonempty
by traversing the correct star-semiring equivalences
(\PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).

@@

\subsection{\GreedySynth}
\label{subsec:greedy-synth}

>>>>

In summary, our main contributions are:
\begin{itemize}
\item We define the space of simple symmetric lenses, which lies between
  asymmetric lenses and full symmetric lenses and has the useful property that
  its members don't have any internal state (\S\ref{sec:overview}). We describe
  a language for simple symmetric string lenses and the cost metric for that
  language (\S\ref{sec:ssl}).
\item We show how to synthesize simple symmetric lenses (including all
  asymmetric lenses as a special case), using a novel application of stochastic
  regular expressions to guide the search process (\S\ref{sec:synthesis}). In
  particular, we show that the RE star semi-ring equivalences remain valid when
  their semantics are enriched by distributions (\S\ref{sec:sre}).
\item We extend the Boomerang implementation with simple symmetric lenses,
  showing that they can easily be combined with a rich variety of other lens
  extensions including quotients and matching lenses. We evaluate our
  implementation over a set of benchmarks drawn from data cleaning, information
  extraction, and file synchronization tasks and demonstrate that we can
  synthesize all of them in under half a minute (\S\ref{sec:evaluation}).
%
  We also analyze the gap between simple and full symmetric lenses
  empirically and argue that simple ones suffice for many applications.
\item We show that the class of simple lenses can be characterized semantically
  as the subset of full symmetric lenses for which a simple ``forgetfulness''
  property holds (\S\ref{sec:relationship}).
\end{itemize}
We finish the paper with a discussion of related work (\S\ref{sec:related}) and
our final thoughts (\S\ref{sec:conc}).

@@

In this section, we give an overview of simple symmetric lenses and our
synthesis algorithm for them, using a simplified example drawn from a hypothetical
U.S. company. In this company, management and human resources (HR) store
information about employees in separate text files: management records the
names of 
employees and their salaries while HR stores the names of employees and their
health insurance providers. Figure~\ref{fig:minimized-representations} gives
examples of the two file formats and regular expressions describing them.

The company uses a simple symmetric lens to keep these files
synchronized. If management adds a new employee, ``Chris Roe: 32500'', a lens
adds the corresponding entry ``Chris Roe, UNK'' to HR's file. The sentinel value
``UNK'' represents the fact that the employee's insurance company is currently
unknown. A similar update happens if HR adds a new employee before management,
in which case the sentinel value ``unk'' reflects the unknown salary information
Furthermore, if HR corrects an error in an employee's name, say changing ``John
Public'' to ``Jon Public'', the lens mirrors that name change in management's
file. Not all the data is mirrored, however. The management file is not updated
in response to insurance changes, and the HR file is oblivious to salary
changes. Simple symmetric lenses are appropriate for keeping these two files
synchronized because the files have a mix of shared and unshared information.

@@

\subsection{Simple Symmetric Lenses}
A simple symmetric lens between formats described by types $X$ and $Y$ is
defined by four functions.\bcp{I think a picture would be helpful. I can
  probably make one pretty easily by tweaking the one from the original
  symmetric paper.} \afm{That would be awesome.}

@@

\begin{gather}
 \tag{1}
 \mbox{$\CreateR \OfType X \rightarrow Y$}\\
 \tag{2}
 \mbox{$\CreateL \OfType Y \rightarrow X$}\\
 \tag{3}
 \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$}\\
 \tag{4}
 \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$}
\end{gather}

\noindent
that collectively satisfy four ``round-tripping'' laws: 
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{gather}
\noindent
The two {\sc Create} functions propagate shared information and fill in default
values when introducing new data. (\EG, the ``unk'' salary entry is added
alongside the name to the management file when HR inserts a new employee). The
two {\sc Put} functions propagate edits from one format to the other by
combining the change from one file with the previous value from the other file.
The roundtripping laws guarantee that if two formats are already synchronized,
{\sc Put}s merely return the second argument, unchanged.

This definition differs from previous formulations of simple symmetric lenses,
as previous formulations required a \emph{complement}. Instead of comprising
four functions, existing formulations merely comprised two functions, which
combined complements with edited formats to create updated complements and
synchronized data. We provide a detailed comparison to classic symmetric lenses
in \S\ref{sec:relationship}.

@@

\subsection{Simple Symmetric Lens Combinators}

Generally, we do not expect programmers to define simple symmetric lenses by
writing the four functions by hand and showing they satisfy the roundtripping
laws. Instead, we provide simple symmetric lens combinators to build up the
requisite lenses from simpler ones, and focus our combinators on providing
lenses between string data formats.

We use regular expressions to denote the types of data formats, and we impose a
typing discipline on our lenses. Specifically, if $\BRegex$ and $\BRegexAlt$ are
regular expressions\footnote{We use \BRegex and \BRegexAlt to denote regular
  expressions. We later use the plain nonterminals \Regex and \RegexAlt for
  stochastic regular expressions.}, then the notation
\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that $\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$. When clear from context, we use the term \textit{lens} to denote a simple symmetric lens.

We briefly explain a selection of these
combinators by using them to define a lens between the two employee data formats
in Figure~\ref{fig:minimized-representations}. The finished lens, which appears
in Figure~\ref{fig:example_lens}, has the type
\lstinline{emp_salaries} $\Leftrightarrow$
\lstinline{emp_insurance}, where
\lstinline{emp_salaries} and
\lstinline{emp_insurance} are regular expressions defined in Figure~\ref{fig:minimized-representations}.

@@

In contrast to the identity lens, the \Disconnect lens does not propagate edits
from one format to the other. The disconnect lens takes four arguments: two
regular expressions and two strings. The regular expressions specify the formats
of the corresponding data on the two sides while the strings provide default
values. On creates, the input values are thrown away, and default values are
returned ($\CreateROf{\String} =
\lstinline{"UNK"}$), and on puts, the second argument is used, and the first is
thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). In our extended
example, the \lstinline{salary} and \lstinline{company} fields are independent,
so we use the disconnect lens to relate them\bcp{Potentially confusing to
  readers (and, even to me, hacky). Why not relate each of them to the constant
  empty string?}\afm{I'm fine with the empty string.  I've seen sentinel values
  like UNK used before in practice, but if it makes the example cleaner and more
  comprehensible, sure.  Anybody else have thoughts on the matter?}.
The defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disconnect(salary, company, "unk", "UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
The insert \lstinline{ins} and delete \lstinline{del} lenses are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source and target formats,
respectively.
\begin{lstlisting}
 ins($t$) = disconnect("", $t$, "", $t$)
 del($s$) = disconnect($s$, "", $s$, "")
\end{lstlisting}

@@

We provide typing rules for each of the simple symmetric combinators, most of
which are syntax directed. However, regular expressions have a number of
equivalences on them, and sometimes we must convert one regular expression type
to another, equivalent, regular expression type. Our typing rules include type
equivalence, so if a lens has type $\Regex \Leftrightarrow \RegexAlt$, then it
is also has type $\Regex' \Leftrightarrow \RegexAlt'$, where $\Regex$ is
equivalent to $\Regex'$, and $\RegexAlt$ is equivalent to $\RegexAlt'$. The
flexibility this type equivalence rule provides is useful when a different
representation of a regular expression would facilitate synthesis.

@@

While our example is relatively simple for demonstration, writing such lenses
for real-world formats is quite difficult, as these formats can be large and
represent data in complex and unintuitive ways. To alleviate the difficulty of
manually writing simple symmetric lenses, we wish to synthesize them instead.
Given a pair of regular expressions and a set of input-output examples, we want
to find a simple symmetric lens that is typed by the regular expression pair and
satisfies the input/output examples. We call such a lens a \emph{satisfying
  lens}. In our running example, we wish to synthesize a lens between
\lstinline{emp_salaries} and \lstinline{emp_insurance}, using as an input-output
example the data in Figure~\ref{fig:minimized-representations}. The algorithm
uses such examples to find suitable defaults for \Disconnect lenses and to
figure out how to match data that could be synchronised in multiple ways.

A challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disconnect(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is another valid lens (where \lstinline{rep($\BRegex$)} denotes some string in the
language of $\BRegex$). Furthermore, adding the input-output example in
Figure~\ref{fig:minimized-representations} 
helps, but only a little: it may, in general, require {\em many} examples to rule
out all possible occurrences of nested \Disconnect lenses,
particularly in complex formats. 
%\ksf{I don't understand the following.  We shouldn't explain the issue in
%terms of \PutR and \PutL as we are focused on the combinators rather
%than the raw functions.}
%The default strings given to \Disconnect merely need to be the strings
%in the input-output examples, and the lens will still satisfy the
%specification.  While the disconnect lens is prohibited by two
%examples, or by a carefully chosen input-output example on \PutR
%or \PutL, this problem continues presenting itself for the sublenses,
%and is highlighted even more on larger formats.
While we could ask programmers to provide detailed and precise specifications
for their synthesis tasks, we would prefer an algorithm that can intelligently
choose a lens despite the underspecification. Instead of merely finding
\emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
likely to please the user.

But what is a ``likely'' satisfying lens? We propose the following
heuristic: a satisfying
lens is more likely if it connects more data between the two
formats, \EG, by using an identity rather than a disconnect lens.  The
likeliest satisfying lenses would connect as much data as
possible. More formally, we define the likelihood of a
satisfying lens as the expected number of bits required to recover
data in one format from synchronized data in the other.  Higher
likelihoods correspond to lower numbers of bits. In more
detail, we say that two strings $s$ and $t$ are \emph{synchronized
according to lens $l$} if $l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$.\footnote{
The notation $l.\PutR$ extracts the \PutR function from simple
symmetric lens $l$; similarly for \PutL.}  We say that we
can \emph{recover} $s$ from $t$ using bits $b$ and lens $l$ if we can
reconstruct $s$ from $t$, $b$, and $l$.  For example, given
the \IdentityLens{} lens, we can recover $s$ from $t$ using no bits
because, in this case, $s$ is just $t$.  In contrast, given
the \Disconnect{} lens, we need enough bits to fully encode $s$ in
order to recover it from $t$ because $t$ contains no useful
information.  This calculation shows that if both \IdentityLens{}
and \Disconnect{} are satisfying lenses, then \IdentityLens{} is more
likely.

The expected number of bits required to recover data is a well-known concept in
information theory: it is the \emph{entropy} of the data. Calculating the
entropy of data requires a probability distribution over the space of possible
values for the data. Specifically, given a set $S$ and a probability
distribution over that set $P \OfType S \rightarrow \Reals$, the entropy of $S$
with probability distribution $P$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$.

In our setting, we already have a way of expressing sets of data:
regular expressions.  What we need is a way to express probability
distributions over that data.
To that end, we adopt \emph{stochastic regular expressions}~\cite{?} (SREs), 
which are regular expressions where each operator is annotated with a
probability.  A stochastic regular expression
specifies both a set of strings and a probability
distribution over it.  From this information, we can calculate the
entropy of the SRE.

This infrastructure gives us what we need to find likely satisfying lenses from
a collection of satisfying lenses. All such lenses must synchronize data between
the same two SREs $\Regex$ and $\RegexAlt$. For any such lens $\Lens$, we can
calculate the expected number of bits required to recover a string $s$ in
$\LanguageOf{\Regex}$ from a synchronized string $t$ in
$\LanguageOf{\RegexAlt}$. This expected number of bits is just the conditional
entropy of $\Regex$ given $\RegexAlt$ and $\Lens$. To ensure symmetry, the
likelihood we assign to $\Lens$ is the sum of the conditional entropy of
$\Regex$ given $\RegexAlt$ and $\Lens$, \bcp{Am I supposed to know what this
  means?} and the conditional entropy of $\RegexAlt$ given $\Regex$ and
$\Lens$.\bcp{Why is this the right (or a sensible) thing to do, besides the fact
  that it ensures symmetry?}

@@

Now that we have a way to calculate likelihood, we need a way to search for
likely lenses. Previous work on synthesis~\cite{?,?,?} has shown that we can use
types to dramatically restrict the search space and improve the effectiveness of
synthesis. In our setting, types are pairs of stochastic regular expressions,
which come with a rich algebraic structure that makes each regular expression
semantically equal to infinitely many others. To tame that search space and
leverage the benefits of type-directed synthesis, we split our algorithm into
two communicating synthesizers following earlier
work~\cite{quotient-lens-synthesis}. The first synthesizer, \RXSearch, navigates
the space of semantically equivalent regular expressions by iteratively applying
rewrite rules that preserve both the semantics and the probability distributions
of those stochastic regular expressions. \RXSearch ranks pairs of stochastic
regular expressions by the number of rewrite rule applications required to
obtain the pair from the stochastic regular expressions given as input. It
passes the pairs off to the second synthesizer, \GreedySynth, in rank order with
the smallest first.

At a high level, \GreedySynth looks for lenses between a given pair of
stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the highest
likelihood by performing a type-directed search. In more detail, \GreedySynth
first normalizes the stochastic regular expressions provided by \RXSearch into
\emph{stochastic DNF regular expressions}. These stochastic DNF regular
expressions are, intuitively, stochastic regular expressions with disjunctions
distributed over concatenations, where associativity information has been
removed. Then, \GreedySynth uses the syntax of these n-ary DNF regular
expressions to find highly normalized lenses, in a form we call \emph{simple
  symmetric n-ary DNF lenses}. These lenses include neither a composition
operator nor a type equivalence rule. These restrictions mean that each simple
symmetric n-ary DNF lens is typed by a single pair of stochastic n-ary DNF
regular expressions and \GreedySynth's search space for a given pair of regular
expressions is finite. The variations that \GreedySynth considers include
choosing between \IdentityLens{} and \Disconnect lenses and deciding how to
match multiple occurrences of the same kind of data, \EG, the two occurrences of
\lstinline{name} in the management and HR data formats (we want to map first
names to first names and last names to last names). Finally, \GreedySynth
converts this lens back into a simple symmetric lens by converting the n-ary
forms into the binary forms provided in the surface language.

@@

\section{Stochastic Regular Expressions}
\label{sec:sre}
To characterize the likely lenses, we must compute the expected number of bits
needed to recover a string in one data source from a synchronized string in the
other data source. To do this, we must first develop a probabilistic langauge
model for our language. We do this with \emph{stochastic regular expressions},
regular expressions annotated with probability information. With this, we can
jointly express a language and a probability distribution over that language. In
the following grammar, $\String$ ranges over strings, and $\Probability$ ranges
over real numbers between 0 and 1, exclusive.

@@

In the stochastic regular expression
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
represents the likelihood the string is in $\Regex_1$. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability how
much more likely having a string involving more than $n$ iterations is than
having a string involving exactly $n$ iterations. The semantics of SREs is
defined by the function $P$ below.

@@

\subsection{Stochastic Regular Expression Equivalences}
$\RXSearch$ needs to be able to find equivalent stochastic regular expressions.
However, existing work does not reason about stochastic regular expression
equivalence. We extend the \emph{star-semiring}~\cite{?} equivalences to
stochastic regular expressions.\bcp{Why these?}

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  The proofs of this theorem, and subsequent theorems in this paper, appears in
  the auxiliary materials accompanying this submission.\bcp{Not in the appendix?}
\end{proof}

\subsection{Stochastic Regular Expression Entropy}
The entropy of a data source $S$ is the expected number of bits required to
describe a random element in $S$ ($-\Sigma_{s\in S}P(s)\Log_2P(s)$). The entropy
of a stochastic regular expression can be computed directly from its syntax,
when the stochastic regular expression is unambiguous and contains no nonempty\bcp{empty?}
subcomponents.
\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{s}$
    & =
    & 1\\

@@

Without unambiguity constraints it is not easy to calculate the entropy. For
example, consider the SRE $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$.
Using the above computation, this SRE has an entropy of $1$, but the entropy
should be $0$.\bcp{This doesn't show that it's not easy---it just shows that
one particular way of computing it is wrong.} Similar issues occur\bcp{Can we at least hint what they are?} when trying to find the entropy when
$\emptyset$ is a subterm, so we do not define the entropy on $\emptyset$.

In our work, we do not find this to be a problem\bcp{Why might one expect
  it to be a problem, rather than just a technicality?}: we require
unambiguous regular 
expressions as input to our synthesis procedure, and we can easily preprocess
emptysets out of the stochastic regular expressions that are themselves nonempty
by traversing the correct star-semiring equivalences
(\EG, \PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).

@@

\subsection{\GreedySynth}
\label{subsec:greedy-synth}
\bcp{The following is all pretty incomprehensibly dense.  A little running example
might help.}
